Asymmetric tail association in extreme climatic events affects population-extinction risk

### 2019-10-23 MainText.Rmd
edited the Intro

# Alternative titles
1) Tail associations in environmental variables and extinction risk
2) The influence of correlations in the extremes on extinction risk

@@

With global climate change, extreme climatic events (ECEs) become more frequent. These ECEs 
could pose severe impact on the resistance or resilience of a population and/ or community level through
life-history traits of the species. Therefore predicting the impact of extreme events 
on population extinction risk is necessery. We here introduce the potential use of "tail association" 
property of two correlated ecological variables in modeling the extinction risk of 
an age-structured single patch model. We also extend this idea 
to density-dependent metapopulation models. For metapopulation models, 
we show extinction risk not only depends on the nature of tail association of 
spatially correlated noise but is also influenced by the inherent dynamics 
of the population-model. To illaborate, we demonstrate left tail dependent noise across spatially correlated patches 
cause higher extinction risk than the right tail dependent noise, when the model show undercompensatory dynamics.
But in overcompensatory regime, the trend is opposite. Our study paves the path for future research considering this tail association instead of simple correlation in population viability analysis.

@@

<!--***DAN: 
1) Shyamolina,  was reminded during my editing of BIVAN that "tail dependence" has 
a specific mathematical definition (now given in the Background section of BIVAN), and that
is not what we mean here when we say "tail dependence". So we cannot go on calling it "tail
dependence" in this manuscript - that would be mike me calling my house "the white house" - 
there already is one of those, and it is much better known than my house, so we cannot
attempt to hijack the name. So can you change terminology throughout, please? Let's refer to
it using terminology such as "associations in the tails" or 
"tail associations". I would like you to provide a 
definition at an appropriate place early in the manuscript that we will call a pair
of positively associated variables "left-tail associated" if their association in their 
left tails is stronger than their association in their right tails (in a sense to be defined
mathematically in Methods, and then in methods we will define things based on cor_l-cor_u).
We call a pair of positively associated variable "right-tail associated" if their association
in their right tails is stronger than their association in their left tails.

2) You need to start getting this paper ready for submission - if you want to send it to 
the Ecosphere special issue (what do you think of that?) then we have to get moving. Please 
let me know soon, after reading about Ecospere on their website. If you think this is a 
good idea, let me know and I will send a summary of the paper to Jon Walter to make sure he is still willing to have us submit the paper to him, and to make sure the paper is close enough
to the theme of the special issue that he is going for, so that he will be willing to see it
submitted there.

3) Please
add the figure you have mentioned in the text but that does not exist. Please run a spell
check. Please go and check every sentence throughout the manuscript for grammatical 
errors. I think you will find some 
and will be able to correct them if you focus on that - there are many, and by correcting them
you will not only make the paper better you will get better at writing grammatically correct 
sentences in English.

4) Assuming you want to go for Ecosphere and Jon is OK with us submitting there, please read the "guidelines for authors" section of the Ecosphere website 
and format the paper for Ecosphere.

5) Please make all the improvements that you can see for yourself
are needed! I suspect if you read the paper carefully now after having worked on something
else for a while you will see many things you can improve!
-->

@@

<!--how common are the extreme events?-->
For past few decades, extreme climatic events (ECEs) appeared more common than 
usual due to anthropogenic climate change, e.g. temperature or precipitation extremes,
heat-waves, wild-fire etc. [@ummenhofer2017extreme]. Frequent extreme events 
can shift temperature or precipitation-mean, 
change variability of the distribution, as well as change its 
symmetry or skewness [@hansen2012perception]. Extreme climatic events can pose severe
impacts at the individual, population, community and ecosystem level [@bragazza2008climatic; 
@gutschick2003extreme; @jiguet2011community; @mckechnie2009climate;
@felton2017integrating]. Therefore, 
understanding the long-term impacts of changes 
in ECEs has been a longstanding interest in ecological research 
[@katz2005statistics; @smith2011ecological; @diez2012will; @moritz2013future; @bailey2016tackling].

<!--How extinction risk is related to extreme events : past studies
consequences of extreme events on population life history, community-->
Earlier studies showed ECEs may affect species abundances (species re-ordering) 
<!--***DAN: Shyamolina, I do not know what the parenthesis refers to

***Shyamolina: species reordering is used to access the change in community 
in response to extreme events or global environmental changes, it relates to change 
in relative abundances of dominant species (mainly). For example, as different species have different 
sensitivity levels to extreme condition (means the extreme threshold is not 
constant for all the species in a community), some years with extreme drought favors say c3 grass 
but in the next year c4 grass could be dominant. Such alteration in community composition has "impact on 
ecosystem function, leading to an extreme ecological response". For detailed overview you may see 
@hoover2014resistance, Jones et al. Ecol. lett. 2017 (doi: 10.1111/ele.12864), 
Smith et al., Ecology 2009 (doi:10.1890/08-1815.1).
Actually LAN K. KNAPP and MELINDA D. SMITH did a lot of work on this to access changes in 
plant community response to extreme events 
(this review article: Felton et al. 2017, Phil. Trans. R. Soc. B 372: 20160142 is a nice summary).
-->
<!--***DAN: Shyamolina, you have obviously done a lot of reading to situate your manuscript
in the appropriate literature, and that is great!-->
through life-history traits, alter community structures through cascading effect on 
higher hierarchical levels and lead to local extinction of 
species [@thomas2004extinction; @edgar2010nino; @moreno2011extreme; @hoover2014resistance]. 
<!--Should I mention idea of match-mismatch hypothesis?-->
For example, a study [@chiu2013impact] showed extreme flood affected 
the annual survival of a prey community that in turn affected predators in 
aquatic as well as adjacent riparian habitats via trophic linkages. Another study by 
@frederiksen2008demographic used $43$ years of ring-recovery data 
to examine the demographic impact of ECEs 
through temporal variability in juvenile, immature and adult survival
of European shags on the Isle of May in eastern Scotland.
A relatively recent study [@mcdermott2017sensitivity] investigated the sensitivity 
of $41$ UK butterfly species to four different ECEs (extreme heat, extreme cold, 
extreme precipitation, drought) over $37$ years at the site level, across 
each species' life-stages. Experimental study showed successive ECEs with short time intervals 
could weaken the resistance of plant communities too [@dreesen2014successive]. 
All these studies indicated that environmental stochasticity (through ECEs) can 
make a population vulnerable to extirpation.
<!--All these studies indicated that environmental stochasticity (through ECEs) 
can lead to demographic stochasticity (through influence on reproductive or 
mortality rate) of different age class of a population making it vulnerable 
to local extinction.-->

<!--Earlier study on metapopulation extinction risk : Large-scale spatial synchrony 
may also be generated by other processes such as migration and predation, but regional
stochasticity in the form of spatially correlated weather conditions is
probably the dominant synchronizing mechanism (ref: Nature1998) but what about taildep. effect on extrisk?.-->
In the metapopulation context, regional stochasticity 
can simultaneously affected multiple habitat patches where spatially correlated weather conditions were 
probably the dominant synchronizing mechanism for extinctions [@hanski1998metapopulation]. A 21 years-long
study by @tack2015increasing on the Glanville fritillary butterfly demonstrated that frequency 
of drought during early larval instars increased amplitude and span of spatial synchrony 
over a metapopulation network of 4000 dry meadows and could affect the long-term population viability.
Therefore, understanding the synchronized fluctuation over a metapopulation-network
remains important to the conservationists because well connected patches, if synchronous, are more 
prone to simultaneous extinction during extreme bad years [@earn2000coherence].

<!--concept for correlation in the extreme tail: tail association-->
Though earlier literature considered the influence of spatial and/ or temporal 
correlations in environmental variables on extinction risk of a metapopulation, to 
our knowledge the separate influence on extinction risk of "tail associations" between
environmental time series in different locations has hardly been considered. "Tail associations" are
the association between two random variables in their extreme values. If two positively correlated 
variables are more associated in their smaller values than larger, they are said to have stronger
"left-tail" or "lower-tail" association (Figure \ref{fig_intro_cop}A, B). Similarly, if two positively correlated variables are more associated in their larger values than smaller, they are 
said to have stronger "right-tail" or "upper-tail" association (Figure \ref{fig_intro_cop}C, D). 
Figure \ref{fig_intro_cop} shows two random variables can have a wide 
range of assciation from extreme to moderate degree in one of their 
extreme tails - often refers to having 'asymmetric tail' - but with a 
fixed Kendall's $\tau$. One can control the degree of "tail asymmetry" - moderate 
to extreme - using "copula", a statistical 
tool generally used to model dependencies between correlated variables (for details see @Ghosh_copula and \nameref{Methods}). 
This 'asymmetric tail' is commonly found in many 
bivariate or multivariate dependence structures for environmental, species-trait, 
phenological, population, community, and ecosystem functioning datasets 
and can help to interpret the causal mechanisms as well as 
to draw consequential inferences for such patterns [@Ghosh_copula]. For example, it 
is well known that spatial dependence of environmental
variables is an important driver of spatial dependence in fluctuating 
population time series (called 'spatial synchrony'). Therefore, one may expect 
the synchronous populations over a larger network would suppress for extreme bad 
years of limiting resources than moderately 'good' years. Any two such populations, that are 
affected badly in a similar manner due to common food shortage, are expected to be more
"left-tail" associated (Figure \ref{fig_intro_cop}A, B) rather than "right-tail" (Figure \ref{fig_intro_cop}C, D) as the environmental fluctuation 
time-series have more "left-tail" association for those extreme bad years
(i.e., 'Moran-effect' like mechanism). Similarly, during 
favorable years if pest metapopulations 
grow abundantly compared to other average years, population time series look like 
\ref{fig_intro_cop}C or D with stronger "right-tail" association, turning local 
outbreaks into regional epidemics. In a 
nutshell, "tail asymmetry" could have profound effect in understanding 
risk for population viability or ecosystem function
considering their relation with extreme climatic events.

A recent study [@Ghosh_copula] used the above mentioned feature of "tail asymmetry" 
to estimate extinction risk for a density-independent metapopulation model. The study 
showed "left-tail" associated density-independent metapopulations 
to be more prone to extinctions than "right-tail" associated one as we may expect species 
are more vulnerable to "catastrophes" (extreme bad years) than to "bonanzas" (extreme good years).
Present manuscript is an extension of that concept in estimating the extinction risk 
beyond the density-independent model. Population growth rate, when depending 
on its own density, serves as a more realistic model to regulate unbounded growth, hence 
models with "density-dependence" are often studied in population ecology to 
predict population viability over long-term period [@hassell1975density; @fowler1987review; 
@courchamp1999inverse]. Here, we will explore the general finding from @Ghosh_copula in the context of
the density-dependent metapopulation patches which experience environmental noises with same spatial correlation but with a variety of tail 
asymmetry (from extreme to moderate, Fig. \ref{fig_intro_cop}). It is expected 
that extreme left-tail associated noise might wipe out some metapopulation patches 
during unfavorable years unless they are rescued by the nearby populations. This scenario 
could be more complex and specific of the direction of association 
depending on the dynamics of the model (see \nameref{Results}).

Tail asymmetry in estimating extinction risk can also be applied to an age structured model. From ecological perspective, it is reasonable to expect that extreme cold winter during growing season
might reduce the adult reproduction rate (fecundity) as well as larval survival. This scenario would resemble Fig.\ref{fig_intro_cop}(A, B) with stronger tail association between vital rates of two stages at their smaller values than larger ones and thereby inflate the population viability compared to a scenario as of Fig.\ref{fig_intro_cop}(C, D). <!--***Shyamolina: Should I tell this? 
It is reasonable to assume cold winter-temperature as the limiting factor 
for each age-structured population - thereby also showing "left-tail" asymmetry between 
the temperature and each age-population time series.(Ref: driver-synchrony relationship as shown in our aphid-plankton paper)--> Though tail association 
property of a copula has been used for risk-analysis in other fields like finance, 
hydrology etc. [@alexander2008market; @li2013; @chen2011flood; @Goswami2018], but to our 
knowledge asymmetric tail association have never been considered 
in estimating the effect of ECEs with age structured 
model or density-dependent metapopulations.

<!--***DAN: Shyamolina, you need to make two major changes in how tail associations
are introdcued. 
  First, do it prior to saying anything about copulas. We have been wanting
to disassociate this paper, to some extent, from BIVAN. Well, probably the best way
to do that is by introducing tail dependence first, before copulas. You can 
describe it intuitively as association in the tails, and you can use pictures to describe it.
You can introduce the partial Spearman correlation (briefly in the Intro, more detail in Methods). 
You will need an explnation of why we 
use Spearman instead of Pearson (talk to me if you do not know, but you may want to read the 
new Background section of BIVAN again first). One does not need to know anything about copulas
to understand tail associations.

  Second, the introduction to copulas that is given is outmoded. See the BAckground section
to BIVAN for improvements. But you do not want to say anything like that much about copulas 
here, you just want to briefly link to the concept, cite some of the Intro work that
has been done (includeing BIVAN and some of the paper BIVAN cites for Intro work),
mention that tail associations are one aspect of copula structure that we have found important,
and introduce any specific things you will need, like the Clayton copula.

Right now this is quite a short paper. You have space to explain things and show pedagogical
figures.
-->

<!--our objective ?-->
In this paper, we will explore the potential use of asymmetric tail association 
for understanding the extinction risk of populations towards extreme climatic event
by addressing the following questions: (Q1) Does nature of tail association 
(left tail or right tail) in vital rates
of a single-patch age-structured population model affect extinction risk of that population differently?
We hypothesized population might have higher extinction risk if vital rates are left-tail dependent.
(Q2) How do asymmetric tail associations in environmental noises across multiple patches influence extinction risk of a density dependent metapopulation model, and how do the effects 
interact with the nature of density dependence? Will they remain consistent with the 
general finding from density-independent metapopulation model [@Ghosh_copula]
(higher extinction risk if noises across patches were left-tail associated than the one 
if noises had right-tail association) irrespective of the nature of density dependence?

@@

<!--FIG : pedagogical fig showing 4 types of tail associations: 
extreme left, moderate left, moderate right, extreme right-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/introfig_copula.pdf}
\caption{Two random variables can have same Kendall 
correlation ($\tau=$ `r ktau_common`), but different degree of associations
in their extreme values: (A) extreme left tail, (B) moderate left tail, (C) moderate right tail, and
(D) extreme right tail association.\label{fig_intro_cop}}
\end{center}
\end{figure*}

>>>>

<!--
Alternative titles
1) Tail associations in environmental variables and extinction risk
2) The influence of correlations in the extremes on extinction risk
DAN FAVORITE SO FAR 3) Tail associations in environmental variables and their impact on extinction risk
-->

@@

<!--***DAN: Dan has not edited this--> 
With global climate change, extreme climatic events (ECEs) become more frequent. These ECEs 
could pose severe impact on the resistance or resilience of a population and/ or community level through
life-history traits of the species. Therefore predicting the impact of extreme events 
on population extinction risk is necessery. We here introduce the potential use of "tail association" 
property of two correlated ecological variables in modeling the extinction risk of 
an age-structured single patch model. We also extend this idea 
to density-dependent metapopulation models. For metapopulation models, 
we show extinction risk not only depends on the nature of tail association of 
spatially correlated noise but is also influenced by the inherent dynamics 
of the population-model. To illaborate, we demonstrate left tail dependent noise across spatially correlated patches 
cause higher extinction risk than the right tail dependent noise, when the model show undercompensatory dynamics.
But in overcompensatory regime, the trend is opposite. Our study paves the path for future research considering this tail association instead of simple correlation in population viability analysis.

@@

<!--***DAN: 
3) Please
add the figure you have mentioned in the text but that does not exist. Please run a spell
check. Please go and check every sentence throughout the manuscript for grammatical 
errors. I think you will find some 
and will be able to correct them if you focus on that - there are many, and by correcting them
you will not only make the paper better you will get better at writing grammatically correct 
sentences in English. 
-->

@@

<!--ECEs have a lot of impacts, it is important to understand their impacts-->
Over the past few decades, extreme climatic events (ECEs) such as
temperature or precipitation extremes, heat-waves, and extreme wildfires
have become more common or more extreme due to anthropogenic climate change 
[@ummenhofer2017extreme]. ECEs are well known to produce severe
impacts at the population, community and ecosystem levels [@bragazza2008climatic; 
@gutschick2003extreme; @jiguet2011community; @mckechnie2009climate;
@felton2017integrating], and thus the goal of improving our understanding of
the variety of ecological impacts of ECEs is an important applied goal for 
modern ecology [@katz2005statistics; @smith2011ecological; @diez2012will; @moritz2013future; @bailey2016tackling]. Frequent extreme events 
can help alter the mean temperature or precipitation of an area, 
can be associated with changes through time in the overall variance of 
climate signals, and can directly contribute to changes in the symmetry or
skewness of distributions of climate signals through time 
[@hansen2012perception]. All such changes have the potential to influence 
populations, including their risk of extirpation from an area 
[@GarciaCarerras2013; @Vasseuretal2014].
For instance, @frederiksen2008demographic used $43$ years of ring-recovery data 
to examine the demographic impact of ECEs on European shags 
via their effects on temporal variability in 
juvenile, immature and adult survival rates.
@mcdermott2017sensitivity investigated the sensitivity 
of $41$ UK butterfly species to four different types of ECEs. 
At the community level, research has showed that ECEs can reorder
the dominance heirarchy of species in an area, sometimes leading to 
selective local extictions, and can also produce effects that cascade 
across trophic levels [@thomas2004extinction; @edgar2010nino; @moreno2011extreme; @hoover2014resistance]. 
For example, @chiu2013impact showed that extreme floods affected 
the annual survival of a prey community, in turn influencing predators in 
aquatic as well as adjacent riparian habitats. 
@dreesen2014successive, in an experimental study, showed that 
successive ECEs with short time intervals between them
can weaken the resistance of plant communities.
<!--***DAN: Shyamolina, what, precisely, does it mean to "weaken the resistance of a
plant community"? Resistance to what? Please clarify.-->

<!--Earlier study on metapopulation extinction risk : Large-scale spatial synchrony 
may also be generated by other processes such as migration and predation, but regional
stochasticity in the form of spatially correlated weather conditions is
probably the dominant synchronizing mechanism (ref: Nature1998) but what about taildep. effect on extrisk?.-->

<!--ECEs have impacts in the spatial context as well-->
ECEs also have impacts in ecological scenarios for which spatial dynamics are
important.
In the metapopulation context, spatial synchrony of population dynamics, which can
be caused by spatial synchrony of climatic fluctuations, is linked to metapopulation
extinction risk [@hanski1998metapopulation; @earn2000coherence].
If ECEs are also spatially more extensive than moderate events,
as seems likely for some types of events, extreme events may accentuate
population synchrony and thereby increase extinction risk.
For instance, a 21-year-long
study of @tack2015increasing on the Glanville fritillary butterfly demonstrated that frequency 
of drought during early larval instars increased the strength and extent 
of spatial synchrony 
over a metapopulation network of 4000 dry meadows, and could thereby influence
the long-term viability of the species in the area of study.

<!--concept for correlation in the extreme tail: tail association-->
Though earlier literature considered the influence of spatial synchrony 
in environmental variables on metapopulation extinction risk, to 
our knowledge the distinct influence of "asymmetric tail associations" between
environmental time series in different locations has hardly been considered. 
"Tail associations" are
associations between two random quantities in their extreme values. 
For instance, if two positively associated 
variables are more associated in their smaller values than in their larger values,
they are said to have stronger
"left-tail" or "lower-tail" association (Figure \ref{fig_intro_cop}A, B). 
If two positively associated variables are more associated in their larger 
values than in their smaller values, they are 
said to have stronger "right-tail" or "upper-tail" association 
(Figure \ref{fig_intro_cop}C, D). 
Figure \ref{fig_intro_cop} shows that two variables can have a wide 
range of patterns of tail association, from extreme left- to extreme right-tail
association, with the same overall correlation. So tail association
is a distinct concept from correlation itself, and may have distinct ecological
ramifications [@Ghosh_copula]. Asymmetric tail associations have been documented 
between environmental variables, and between the same environmental variable 
measured in different places [@Serinaldi2008; @li2013; @She2018; @Goswami2018]. 
Asymmetric tail associations have also been documented between numerous ecological
variables and between environmental and ecological variables [@Ghosh_copula].

<!--ECEs probably produce tail associations-->
There are reasons to believe that ECEs may produce asymmetric tail 
associations between climatic variables
measured in different locations. For instance, if a single ECE is associated
with spatially widespread, large values 
of some climatic variable, but small values of the
variable are associated with local weather rather than ECEs, the result could
be stronger right-tail association between measurements of the climatic variable
made in different locations, through time.
For example, rain storms which deposit exceptionally large amounts of rain
may do so over large areas; but months with less rain may experience
scattered showers only, which may be much less correlated across locales.

<!--***DAN: May be scope for a discussion para linking what is known about
how ECEs are changing to what effect that may have on how tail assiciations
in environmental variables might be changing, and what influence that may have
on extinction risk. If little is known, then this might be a good topic for 
future work.-->

There are also reasons to hypothesize that asymmetry of tail association in 
environmental variables may influence extinction risks for metapopulations.
For instance, assuming for the sake of a simple example that low values of an environmental
variable are "bad" for populations of a given species and high values are "good",
greater left-tail associations between measurements of the variable in different
habitat patches may cause higher metapopulation extinction risk because then
very bad years for component populations would occur at the same time in 
many patches, reducing the potential for rescue effects. Greater left-tail associations
would probably not have the same effect, even if overall correlations between locations
were the same, because it that case it would be very good years for the component 
populations which would occur at the same time in many patches [@Ghosh_copulas].  
@Ghosh_copula substantiated this intuition by showing that it held true for a 
modelling setup using a spatial version of a very simple non-density dependent
population growth model (the Lewontin-Cohen model). In essence, if environmental
"catastrophes" (i.e., extremely bad years for a population) are widely 
spatially synchronized it should create much greater metapopulation extinction 
risk than if "bonanzas" (i.e., extremely good years) are widely synchronized, even
if overall (non-tail-specific) levels of environmental synchrony are the same
in both these cases. The present study extends tests of this idea to metapopulation
models incorporating density dependence.

Extinction risk is also very commonly estimated for age-structured models 
such as stochastic matrix models [@MorrisDoak2002 ; @Caswell2006] and 
integral projection models
[@Merowetal2014]. For such models, extinction risk depends, in part, on the long-term
population growth rate, a population parameter which in turn 
depends on the means, variances and covariances of population vital rates such 
as the survival and fecundity rates of age classes. Extinction risks may,
*a priori*, also depend on possible asymmetries of tail association between
population vital rates. For instance, it seems reasonable to imagine
a case in which both adult fecundity and juvenile survival rates 
depend on an environmental variable such as spring rainfall in similar ways, but 
only when that variable takes very low values. When rainfall takes
typical or high values, fecundity and survival may be limited instead 
by different factors. This could produce left-tail dependence between these 
vital rates, possibly resulting in different extinction risk for the model
compared to a case in which vital rates were symmetrically associated in 
their tails but otherwise statistically similar.

<!--objectives, last para of intro-->
In this paper, we will explore the potential for asymmtric tail associations to 
influence extinction risk in populations and metapopulations. Specifically,
we will address the following sets of questions. Q1) Can the nature of tail associations
between vital rates in an age-structured population model influence
extinction risk for the model, and to what extent? Q2) Does the nature
of tail associations between measurements of an environmental variable in different
locations influence extinction risk of a density dependendent metapopulation model
that depends on the environmental variable? How does the nature of the density
dependence mediate this influence? Will density dependence reverse or reproduce
the earlier result [@Ghosh_copula], using non-density-dependent models, that 
left-tail dependence in environmental variables accentuates and right-tail dependence 
mitigates extinction risks?
For Q1, we use a simple 2-stage model. We hypothesize that left-tail dependence
between juvenile survival and adult fecundity rates increases extinction risk
relative to right-tail dependence, even when overall correlation is kept fixed. 
For Q2, we consider spatial extensions of several simple and commonly
used population models. Though tail associations are used
to assess risks in fields such as finance and hydrology 
[@alexander2008market; @li2013; @chen2011flood; @Goswami2018], to our
knowledge asymmetric tail associations have rarely been assessed for their
importance for extinction risks of populations and species (but see @Ghosh_copulas). 
Our questions are an important step for understanding
the potential influence of ECEs on extinction risks if ECEs are associated with
tail dependence in environmental variables.

@@

<!--FIG : pedagogical fig showing 4 types of tail associations: 
extreme left, moderate left, moderate right, extreme right-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=15 cm]{./Results/introfig_copula.pdf}
\caption{Two stochastic quantities can have same Kendall 
correlation, but different degrees of tail association: 
(A) extreme left-tail association, (B) moderate left-tail association, (C) 
moderate right-tail association, and (D) extreme right-tail association.
In all cases Kendall's $\tau$ was the same ($\tau \approx$ `r ktau_common`),
up to sample variation (the sample correlation for each dataset is displayed
on its panel). The extreme left-tail (respectively, right-tail) dependent 
case is perfectly correlated below (respectively, above) the threshold of 0.
See Methods for information on how bivariate noise was generated with
asymmetric tail associations.
\label{fig_intro_cop}}
\end{center}
\end{figure*}
<!--***DAN: Shyamolina, can you please put the actual Kendall tau that pertained
for each dataset on each corresponding
panel? Use, say, 3-4 digits accuracy for display. These values should be very similar, 
but are expected to differ due to sampling variation. I would like to make that
clear.-->
